---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE
)
```

# Cubble

<!-- badges: start -->
[![R-CMD-check](https://github.com/huizezhang-sherry/cubble/workflows/R-CMD-check/badge.svg)](https://github.com/huizezhang-sherry/cubble/actions)
<!-- badges: end -->

Cubble is a vector spatio-temporal data structure for data analysis and visualisation.

## Installation

You can install the released version of cubble from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cubble")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("huizezhang-sherry/cubble")
```
## Example

Using `as_cubble()` to create a cubble in the nested form by supplying the site identifier `key`, temporal identifier `index`, and the spatial coordinates that defines the site `coords`.

```{r}
library(cubble)
library(dplyr)
climate_flat %>% 
  as_cubble(key = id, index = date, coords = c(long, lat))
```

Use `stretch()` to switch from the nested form to the long form. Long form is convenient for time-wise computation, for exmaple, we can filter to January records:

```{r}
climate_flat %>% 
  as_cubble(key = id, index = date, coords = c(long, lat)) %>% 
  stretch() %>% 
  filter(lubridate::month(date) == 1)
```

Switch back to the list-column form with `tamp()`. List-column form is suitable for site-wise manipulation, for example, we can add the count of no-raining days for each station: 

```{r}
climate_flat %>% 
  as_cubble(key = id, index = date, coords = c(long, lat)) %>% 
  stretch() %>% 
  filter(lubridate::month(date) == 1) %>% 
  tamp() %>% 
  mutate(zero_rain = sum(ts$prcp == 0, na.rm = TRUE))
```

