---
title: "app-climate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{app-climate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
```

```{r setup}
library(cubble)
library(stringr)
library(ggplot2)
library(fabletools)
library(tidyr)
library(broom)
library(dplyr)
library(feasts)
library(rmapshaper)
library(forcats)
library(lubridate)
state_map <- ms_simplify(ozmaps::abs_ste, keep = 2e-3)
filter <- dplyr::filter
```

```{r}
df <- oz_climate %>% 
  mutate(name = as.factor(str_to_lower(str_remove(name, ", AS")))) %>% 
  as_tsibble(index = date, key = station) %>% 
  global(station) 
  
#  rename(elev = elevation)
# add rename: https://github.com/tidyverse/dplyr/blob/master/R/rename.R

df_daily <- df %>% 
  mutate(tmax_missing = all(is.na(ts$tmax)),
         tmin_missing = all(is.na(ts$tmin))) %>% 
  filter(!tmax_missing, !tmin_missing, 
         name != "coffs harbour mo",# too many missing from 2016 - 2020 for coffs harbour mo
         station != "ASN00016098" # almost all missing for tmin
         ) 

df_daily %>% 
  zoom() %>% 
  migrate(name) %>% 
  ggplot(aes(x = date ,y = tmax)) + 
  geom_line() + 
  facet_wrap(vars(name))

```

## Location of all the stations

```{r}
ggplot() +  
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(data = df, aes(x = long, y = lat)) + 
  coord_sf() + 
  theme_void()
```

## Features on tmax

```{r}
feat_tmax <- df_daily %>% 
  tidyr::unnest(ts) %>% 
  as_tsibble(index = date, key = station) %>% 
  features(tmax, feature_set(pkgs = "feasts"))

feat_tmax_clean <- feat_tmax %>% 
  dplyr::select(-contains(c("lb", "bp", "zero", "pp", "diffs", "lambda")))

pca_res <- feat_tmax_clean %>% 
  dplyr::select(-station) %>% 
  prcomp(scale = TRUE) 

pca_tmax <- pca_res %>% 
  augment(feat_tmax_clean)
```

```{r}
df_tmax_pca <- df_daily %>% 
  left_join(pca_tmax %>% dplyr::select(station, .fittedPC1, .fittedPC2)) 

df_tmax_pca %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2)) + 
  geom_point() + 
  coord_equal()
```

think about how to show these three stations are different from others

```{r}
# see those stations on the map
ggplot() +  
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(data = df, aes(x = long, y = lat)) + 
  geom_point(data = df_tmax_pca %>% filter(.fittedPC1 < -5), 
             aes(x = long, y = lat), color = "red") + 
  coord_sf() + 
  theme_void()

# average across station doesn't work currently :(
average <- df_tmax_pca %>% 
  filter(.fittedPC1 >= -5) %>% 
  zoom() %>%
  as_tibble() %>% 
  group_by(date) %>% 
  summarise(tmax = mean(tmax, na.rm = TRUE)) %>% 
  mutate(name = "average")

# plot the tmax series along with the average of all others - averaging makes the series smoother ....
df_tmax_pca %>% 
  filter(.fittedPC1 < -5) %>% 
  zoom() %>% 
  migrate(name) %>% 
  bind_rows(average) %>% 
  ggplot(aes(x = date , y = tmax)) + 
  geom_line() + 
  facet_wrap(vars(name), ncol = 1)

```

## Features on prcp

```{r}
df_prcp <- df %>% 
  mutate(prcp_missing = all(is.na(ts$prcp) | ts$prcp == 0)) %>% 
  filter(!prcp_missing) 

feat_prcp <- df_prcp %>% 
  tidyr::unnest(ts) %>% 
  as_tsibble(index = date, key = station) %>% 
  features(prcp, feature_set(pkgs = "feasts"))

feat_prcp_clean <- feat_prcp %>% 
  dplyr::select(-contains(c("lb", "bp", "zero", "pp", "diffs", "lambda", "pacf", "kpss", "hurst")))

pca_prcp <- feat_prcp_clean %>% 
  dplyr::select(-station) %>% 
  prcomp(scale = TRUE) %>% 
  augment(feat_prcp_clean)
```

```{r}
df_prcp_pca <- df_prcp %>% 
  left_join(pca_prcp %>% dplyr::select(station, .fittedPC1, .fittedPC2)) 

df_prcp_pca %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2)) + 
  geom_point() + 
  coord_equal()

```

```{r}
# see those stations on the map
ggplot() +  
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(data = df, aes(x = long, y = lat)) + 
  geom_point(data = df_prcp_pca %>% filter(.fittedPC1 > 4), 
             aes(x = long, y = lat), color = "red") + 
  coord_sf() + 
  theme_minimal()

# plot the tmax series along with the average of all others
df_prcp_pca %>% 
  filter(.fittedPC1 > 4) %>% 
  zoom() %>% 
  migrate(name) %>% 
  ggplot(aes(x = date , y = prcp)) + 
  geom_line() + 
  facet_wrap(vars(name))

```


## Features on tmin

```{r}
df_tmin <- df %>% 
  mutate(tmin_missing = all(is.na(ts$tmin)), 
         first_day = is.na(ts$tmin[1])) %>% 
  filter(!tmin_missing, !first_day) 

feat_tmin <- df_tmin %>% 
  tidyr::unnest(ts) %>% 
  as_tsibble(index = date, key = station) %>% 
  features(tmin, feature_set(pkgs = "feasts"))

feat_tmin_clean <- feat_tmin %>% 
  dplyr::select(-contains(c("lb", "bp", "zero", "pp", "diffs", "lambda", "pacf", "hurst", "kpss", "spikiness"))) %>% 
  filter(station != "ASN00040223")

pca_tmin <- feat_tmin_clean %>% 
  dplyr::select(-station) %>% 
  prcomp(scale = TRUE) %>% 
  augment(feat_tmin_clean)
```

```{r}
df_tmin_pca <- df_tmin %>% 
  left_join(pca_tmin %>% dplyr::select(station, .fittedPC1, .fittedPC2)) 

df_tmin_pca %>% 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2)) + 
  geom_point() + 
  coord_equal()
```

```{r}
# see those stations on the map
ggplot() +  
  geom_sf(data = ozmaps::abs_ste, aes(geometry = geometry)) +
  geom_point(data = df, aes(x = long, y = lat)) + 
  geom_point(data = df_tmin_pca %>% filter(.fittedPC1 > 6), 
             aes(x = long, y = lat), color = "red") + 
  coord_sf() + 
  theme_minimal()
```


## Cross linking on map and PC plot

```{r}
library(crosstalk)
library(plotly)
shared_dt <- SharedData$new(df_tmax_pca %>% mutate(less_than_5 = ifelse(.fittedPC1 < -5, TRUE, FALSE)))
p1 <- shared_dt %>%  
  ggplot(aes(x = .fittedPC1, y = .fittedPC2)) + 
  geom_point() + 
  coord_equal()

p2 <- ggplot() +  
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(data = shared_dt, aes(x = long, y = lat), color = "red") + 
  coord_sf() + 
  theme_void()

bscols(ggplotly(p1), ggplotly(p2))
```



## Decomposition

```{r}
df_monthly <- df_daily %>% 
  zoom() %>% 
  mutate(ym = tsibble::yearmonth(date)) %>% 
  index_by(ym) %>% 
  summarise(tmax = mean(tmax, na.rm = TRUE),
            tmin = mean(tmin, na.rm = TRUE), 
            prcp = sum(prcp, na.rm = TRUE)) %>% 
  ungroup(ym) %>% 
  filter(year(ym) < 2021) %>% 
  mutate(year = as.factor(year(ym)),
         month = as.factor(month(ym)), 
         prcp = sqrt(prcp + 0.001))
```


Link monthly tmax with map


```{r}
smooth_plot <- df_monthly %>% 
  ggplot(aes(x = as.Date(ym), y = tmax, group = station)) + 
  stat_smooth(se = FALSE) # gam

smoother <- layer_data(smooth_plot) %>% 
  group_by(group) %>% 
  summarise(y_diff = max(y) - min(y))

df_tmax_smoother <- df_monthly %>% 
  global() %>% 
  bind_cols(smoother) %>% 
  arrange(-y_diff)

shared_dt <- df_tmax_smoother %>% 
  zoom() %>% 
  migrate(lat, long, y_diff) %>% 
  highlight_key(~station)

p1 <- shared_dt %>% 
  ggplot() + 
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(aes(x = long, y = lat, color = y_diff, group = station)) + 
  theme_void()

map <- highlight(ggplotly(p1), selectize = TRUE, on = "plotly_selected")

p2 <- shared_dt %>% 
  ggplot(aes(x = as.Date(ym), y = tmax, group = station, color = y_diff), alpha = .1) + 
  geom_line()

line <- highlight(ggplotly(p2), on = "plotly_hover")

bscols(map, line)

```

See yearly variation of different stations

```{r}
df_into_month <- df_monthly %>% 
  group_by(month, station) %>% 
  summarise(tmax = mean(tmax, na.rm = TRUE),
            tmin = mean(tmin, na.rm = TRUE)) %>% 
  ungroup() %>% 
  left_join(meta(df_monthly)) %>% # a bit of hacky to get around with group_by
  global(station)
```

tmax

```{r}
 df_monthly %>% 
    ggplot(aes(x = month, y = tmax, group = interaction(year, station), color = year)) + 
    geom_line() + 
    facet_wrap(vars(station))
```

Temperature difference for stations

Summarises the monthly mean temperature across each year for all the stations and calculate the mean temperature difference. Plot the two averaged temperature series as line and on the map for each station. 

Those stations along the coastline are the ones with small temperature difference and this makes sense because water body has the effect to neutralise temperature. 

TODO: Add linking 

```{r}
df_monthly_ordered <- df_into_month %>% 
  mutate(tdiff = mean(ts$tmax - ts$tmin, na.rm = TRUE)) %>% 
  zoom() %>%
  migrate(tdiff, name)

p1 <- df_monthly_ordered %>% 
  ggplot() + 
  geom_line(aes(x = month, y = tmin, group = name, color = tdiff)) + 
  geom_line(aes(x = month, y = tmax, group = name, color = tdiff)) + 
  facet_wrap(vars(fct_reorder(name, -tdiff)), labeller = label_wrap_gen(width=20)) + 
  ylab("temperature")

p2 <- df_monthly_ordered %>% 
  global() %>% 
  ggplot() + 
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(aes(x = long, y = lat, color = tdiff, group = name)) + 
  theme_void()

p1 | p2
```

prcp

- steps

Those with more rains are in the north or along the eastern coast

- add interactivity
- investigate station ASN00090015

```{r}
df_prcp_weekly <- df_daily %>% 
  zoom() %>% 
  mutate(week = week(date)) %>% 
  index_by(week) %>% 
  summarise(prcp = sum(prcp, na.rm = TRUE)) %>% 
  ungroup(week) %>% 
  mutate(prcp = sqrt(prcp + 0.001)) %>% 
   global() %>% 
  mutate(prcp_max = max(ts$prcp))

p1 <- df_prcp_weekly %>% 
  zoom() %>% 
  migrate(prcp_max) %>% 
  ggplot(aes(x = week, y = prcp, group = station, color = prcp_max)) + 
  geom_line() + 
  facet_wrap(vars(station))

p2 <- df_prcp_weekly %>% 
  ggplot() + 
  geom_sf(data = state_map, aes(geometry = geometry)) +
  geom_point(aes(x = long, y = lat, group = name, color = prcp_max)) + 
  theme_void()

p1 | p2
```

see also how those heavily rained stations map to the geo map

also the line plot looks spike - try with weekly aggregation rather than monthly?
  
  
  
  also use glyph map
