---
title: "0. Import data as a cubble"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{import}
  %\VignetteEngine{knitr::rmarkdown}
   %\usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE,
  message = FALSE
)

library(cubble)
library(dplyr)
```

This article shows you how to create a cubble from data in the wild. You should have already seen an example of constructing a cubble from a tibble in the README page and here are more examples that construct a cubble from:  

  - a tibble with list column  
  - a tsibble 
  - separate spatial and temporal tables
  - NetCDF data
  - a `stars` object
  
## A cubble from tibble with list column

List column can be a useful structure when querying data since you can first create a row-wise data frame with all the metadata and then query the data as a nested list. We will show how to convert this output into a cubble in a small example with five weather stations close to Sydney, Australia. Here is the metadata of the stations:

```{r echo = FALSE}
syd_all <- climate_aus %>% slice_nearby(coord = c(151.207, -33.867), n = 5)
syd <- syd_all %>% select(id:wmo_id)
```

```{r}
syd
```

Temporal variables can be queried by `rnoaa::meteo_pull_monitors()` with the date range and variable supplied. Here we turn the `stations` data into a rowwise data frame and then query the climate variables as a list-column, `ts`:  

```{r eval = FALSE}
raw <- syd %>%  
  rowwise() %>%  
  mutate(ts = list(rnoaa::meteo_pull_monitors(id, 
                                       date_min = "2020-01-01", 
                                       date_max = "2020-12-31",
                                       var = c("PRCP", "TMAX", "TMIN")) %>%  select(-id))) 
raw
```


```{r echo = FALSE}
raw <- syd_all %>% select(-c(long_ref, lat_ref, dist)) %>% rowwise()
raw
```


A cubble can then be created by supplying:

  - `id` as the key to identify each station, 
  - `date` as the index to identify time, and 
  - `coords` to identify the spatial coordinates of each station.

```{r}
syd_climate <- raw %>%  
  as_cubble(key = id, index = date, coords = c(long, lat))

syd_climate
```

# A cubble from tsibble

If you have already got a `tsibble` object, with key and index registered, only `coords` needs to be specified to create a cubble:

```{r echo = FALSE}
dt <- climate_flat %>%  
  tsibble::as_tsibble(key = id, index = date)
```

```{r}
dt
dt %>%  as_cubble(coords = c(long, lat))
```

Notice here each element of the list-column `ts` is of class `tbl_ts` and you are free to apply your favourite function on the tsibble class :)

# Separate spatial and temporal tables

Sometimes, you may get the spatial and temporal data from different sources and they live in two separate tables with a linking variable:

```{r}
# a spatial sheet
cubble::stations

# a temporal sheet
cubble::climate
```

When created out of separate tables, `as_cubble()` will check on the match of linking variable in the spatial and temporal table. 

```{r}
as_cubble(list(spatial = cubble::stations, temporal = cubble::climate),
          key = id, index = date, coords = c(long, lat))
```

Messages will emit to notify the unmatched, if detected. This would be useful to catch some sites with only spatial/ temporal information and the slight mismatch in character linking variable.


# A cubble from NetCDF data

NetCDF (Network Common Data Form) is a commonly used data format in the climatology community to deliver global mapping of atmosphere, ocean, and land. Here we provide some general information regarding this format before showing an example of creating a cubble out of NetCDF data. 

NetCDF data has two main components: 

  - **dimension**: define the spatio-temporal grid (longitude, latitude, and time) and 
  - **variable**: populate the defined grid. 
  
Attributes are usually associated with dimension and variable in the NetCDF format data. A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: `ncdf4`, a low-level interface that calls C interface: `RNetCDF`, and a tidyverse implementation: `tidync`. Here let's take a look at a NetCDF data: 

```{r}
path <- system.file("ncdf/era5-pressure.nc", package = "cubble")
raw <- ncdf4::nc_open(path)
raw
```

In a NetCDF data, it is not the the actual data that gets directly printed out, but the metadata. There are 2 variables and 3 dimensions in this data and each is associated with a few attributes. In this data the attributes for the two variables includes the scaling and offset parameter, representation of the missing and its fill value, along with its unit and names. In NetCDF, data is stored in its *packed value* to save space and sometimes, you will need to use a formula like $\text{unpacked value} = \text{packed value} \times \text{scale factor} + \text{add offset}$ to unpack the data. Luckily, when reading in the NetCDF data with the `ncdf4` package, it has already unpack the data for you, so no need to worry about the scaling and offset. 

In principle, NetCDF data can store data with arbitrary variable, dimension, and attribute and this will cause a chaos to generalise its manipulation. [Metadata convention for climate and forecast (CF convention)](http://cfconventions.org/) is a guideline that has been designed to standardise the format of NetCDF data. Thanks to the CF convention, `cubble` can now extract specific components as per the CF convention to build a cubble from NetCDF data. 

Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a cubble. It maps each combination of longitude and latitude into an `id` as the key:

```{r}
dt <- as_cubble(raw, vars = c("q", "z"))
dt
```

The memory limit with NetCDF data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year. You can reduce the spatial grid points in exchange for longer time period and more variables. A 300 by 300 spatial grid can be:
  
  - a bounding box of [100, -80, 180, 0] at 0.25 degree resolution, or 
  - a global bounding box [-180, -90, 180, -90] at 1 degree resolution. 
  
Subsetting longitude and latitude grid is available through `long_range` and `lat_range` if the NetCDF file has finer resolution than needed.

```{r}
dt <- as_cubble(raw, vars = c("q", "z"),
                long_range = seq(-180, 180, 1),
                lat_rnage = seq(-90, -5, 1))
dt
```


# Convert from a `stars` object to `cubble`

Currently, the `cubble` package can convert a `stars` object with three dimension(s) in the order of longitude, latitude, and time: 
  
```{r}
# create a toy stars object
m <- array(1:60, dim = c(x= 5, y = 4, t = 3))
time = 1:3
library(units)
units(time) = as_units("days since 2015-01-01")
m_dim <- stars::st_dimensions(x =  seq(146, 162, 4), y = seq(-44, -41, 1), t = time)
st <- stars::st_as_stars(list(m = m, m2 = m), dimensions = m_dim)
st
as_cubble(st)
```

When the `dimensions` object is too complex for `cubble` to handle, the package will emit an message. 

